---
title: "Linear Regression Mini-competition"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(stats)
library(readxl)
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(tidyverse)
library(GGally)

```
**Load the data**
__Uploaded the data in csv format__

```{r load_Data}
#load data
df_2019 <- read.csv("k12_2019.csv")
df_2016 <- read.csv("k12_2016.csv")
head(df_2019, 5)
```
**Start exploring the data**

```{r}
columns = names(df_2019)
columns
```

```{r visualize_data}
# Subset the columns of interest from df_2019
columns_of_interest <- df_2019[, c("SEGRADES", "CHLDNT", "RACEETH", "P1SEX", "PARGRADEX", "DSBLTY", "TTLHHINC", "EINTNET")]

# Create line graphs for each variable against SEGRADES
line_chldnt <- ggplot(columns_of_interest, aes(x = CHLDNT, y = SEGRADES)) +
  geom_line() +
  labs(x = "CHLDNT", y = "SEGRADES") +
  ggtitle("Relationship between CHLDNT and SEGRADES")

line_raceeth <- ggplot(columns_of_interest, aes(x = RACEETH, y = SEGRADES)) +
  geom_line() +
  labs(x = "RACEETH", y = "SEGRADES") +
  ggtitle("Relationship between RACEETH and SEGRADES")

line_p1sex <- ggplot(columns_of_interest, aes(x = P1SEX, y = SEGRADES)) +
  geom_line() +
  labs(x = "P1SEX", y = "SEGRADES") +
  ggtitle("Relationship between P1SEX and SEGRADES")

line_pargradex <- ggplot(columns_of_interest, aes(x = PARGRADEX, y = SEGRADES)) +
  geom_line() +
  labs(x = "PARGRADEX", y = "SEGRADES") +
  ggtitle("Relationship between PARGRADEX and SEGRADES")

line_dsblty <- ggplot(columns_of_interest, aes(x = DSBLTY, y = SEGRADES)) +
  geom_line() +
  labs(x = "DSBLTY", y = "SEGRADES") +
  ggtitle("Relationship between DSBLTY and SEGRADES")

line_ttlhhinc <- ggplot(columns_of_interest, aes(x = TTLHHINC, y = SEGRADES)) +
  geom_line() +
  labs(x = "TTLHHINC", y = "SEGRADES") +
  ggtitle("Relationship between TTLHHINC and SEGRADES")

# Display the line graphs for each variable against the dependent
#line_chldnt
#line_raceeth
#line_p1sex
#line_pargradex
#line_dsblty
line_ttlhhinc



```

```{r simple_model}
#write the multiple linear regression model
#the dependent variable in this case: SEGRADES
#Using simple ML mol
model <- lm(SEGRADES ~ CHLDNT + RACEETH + P1SEX + PARGRADEX + DSBLTY + TTLHHINC, data = columns_of_interest)

# Print the summary of the linear regression model
summary(model)

```
```{r}
data <- columns_of_interest
# review any visual patterns
  data %>% 
    select(SEGRADES, CHLDNT, RACEETH, P1SEX, PARGRADEX, DSBLTY, TTLHHINC,EINTNET) %>% 
    ggpairs()

  #fit the mlr model
  lm_spec <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")

  mlr_mod <- lm_spec %>% 
  fit(SEGRADES ~ CHLDNT + RACEETH + P1SEX + PARGRADEX + DSBLTY + TTLHHINC + EINTNET, data = data)

  tidy(mlr_mod)
```

**Brief analysis from the ggpairs() above**
1. There was no correlation between the predictors, meaning they can be used within the model in determining the students performance (grades during a school year). 
2. CHLDNT (internet usage by a child in learning activities), DSBLTY (Child with disability) and P1SEX (gender of parent at home) p-values do not show any significance in predicting the child performance from virtual learning. 
3. EINTNET, TTLHHINC, PARGRADEX and RACEETH were therefore suitable in predicting the students who most benefited from virtual learning. 

4. Most students were white non-hispanic, then hispanic, black non-hispanic, and asian.
5.Higher education of parents or gardiuns (Vocational/technical education after high school or some college, College graduate, Graduate or professional school) benefited more.
6. Children from families with high total household income ($75,001 to $100,000, $100,001 to $150,000, $150,001 to $200,000) benefited more.

7. Majority of the students who benefited did not have online classes. This means, most of the successful child's parents prefered child interaction with teachers. 


```{r relationship_among_vars}
#selected model
data %>% 
    select(SEGRADES, RACEETH, PARGRADEX, TTLHHINC,EINTNET) %>% 
    ggpairs()

  #fit the mlr model
  lm_spec_1 <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")

  mlr_mod1 <- lm_spec_1 %>% 
  fit(SEGRADES ~ RACEETH + PARGRADEX + TTLHHINC + EINTNET, data = data)

  tidy(mlr_mod1)

```




